---
title: "Peer Assignment Week 2"
author: "Nate Foulkes"
date: "8/1/2021"
output: html_document
---

## Loading and pre-processing the data
```{r setup, echo=TRUE}
library(dplyr)
library(ggplot2)
if(!file.exists("data")){
        dir.create("data")
}
url1 <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(url1, destfile = "./data/df1.zip")
unzip(zipfile = "./data/df1.zip")
activity <- read.csv("./activity.csv",stringsAsFactors=TRUE, na.strings = "NA")
good <- complete.cases(activity)
no_NA_data <- activity[good,]
activity$date <- as.POSIXlt(activity$date)
```

## What is mean total number of steps taken per day?
```{r, echo=TRUE}
daily_steps <- no_NA_data %>% 
        group_by(date) %>% 
        summarise(TotalSteps = sum(steps))
mean_steps <- mean(daily_steps$TotalSteps)
median_steps <- median(daily_steps$TotalSteps)
```
### Histogram of the daily steps
```{r, echo=TRUE}
hist(daily_steps$TotalSteps , 
     breaks = 10, 
     col = "lightblue", 
     main = "Total Daily Steps", 
     xlab = "steps")
```
### The mean and median number of daily steps
```{r, echo=TRUE}
mean_steps_1 <- mean(daily_steps$TotalSteps)
median_steps_1 <- median(daily_steps$TotalSteps)
```
The mean number of steps `r mean_steps_1`. 
The median number of steps `r median_steps_1`.

# What is the average daily activity pattern?
```{r, echo=TRUE}
avg_step_interval <- no_NA_data %>% group_by(interval) %>% summarise(avgSteps = mean(steps))
plot(avg_step_interval,
     type = "l",
     col = "blue",
     main = "Average Steps by 5-minute Interval",
     ylab = "Average Steps",
     xlab = "Interval")
Five_min_interval <- avg_step_interval$interval[[grep(max(avg_step_interval$avgSteps),avg_step_interval$avgSteps)]]

```
At `r Five_min_interval` is the interval of the maximum number of daily steps.

# Imputing missing values
```{r, echo=TRUE}
my_NAs <- colSums(is.na(activity))
```
The number of rows with NAs is
`r my_NAs`

## Fill in the NA values
### The NA values will be replaced by the mean of that 5-minute interval
```{r, echo=TRUE}
interval_mean <- no_NA_data %>% 
        group_by(interval) %>% 
        summarise(mean = mean(steps))

my_merge <- merge(activity, interval_mean, by = "interval")

```
### New data set with the NA values replaced
```{r, echo=TRUE}
clean_data <- within(my_merge, steps[is.na(steps)] <- mean[is.na(steps)])

plotting_data <- clean_data %>% 
        group_by(date) %>% 
        summarise(TotalSteps = sum(steps))

```

### Histogram of the daily steps
```{r, echo=TRUE}
hist(plotting_data$TotalSteps , 
     breaks = 10, 
     col = "light pink", 
     main = "Total Daily Steps", 
     xlab = "steps")
```

### The mean and median number of steps
```{r, echo=TRUE}
mean_steps_2 <- mean(plotting_data$TotalSteps)
median_steps_2 <- median(plotting_data$TotalSteps)
```

The mean number of steps `r mean_steps_2`.
The median number of steps `r median_steps_2`.

# Are there difference in activity patterns between weekdays and weekends?

## Create a new factor variable for weekdays and weekends
```{r, echo=TRUE}
DoW_df <- mutate(clean_data, day_of_the_week = weekdays(date))

weekend <- c("Saturday", "Sunday")

weekend_df <- DoW_df[DoW_df$day_of_the_week %in% weekend,] %>% 
        group_by(interval) %>% 
        summarise(Avg_step = mean(steps))

weekday <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

weekday_df <- DoW_df[DoW_df$day_of_the_week %in% weekday,] %>% 
        group_by(interval) %>% 
        summarise(Avg_step = mean(steps))
```

## A panel plot to display the time series difference between weekday and weekend
```{r, echo=TRUE}
par(mfrow = c(2,1))
plot(weekday_df, type = "l", main = "Weekday", col = "blue")
plot(weekend_df, type = "l", main = "Weekend", col = "green")
```